---
title: "Library size"
output: html_document

---


## Quality control


The preprocessed samples were ordered based on the library size.
In this stage, samples with less than `r min_reads` reads were filtered out.
Any Negative Control that has 0 reads would also be removed.


```{r librarySize,  message = FALSE, fig.cap = "Reads library size after preprocessing"}

## Filter out Mocks
#expReads <- sampTrack %>% filter(Sample_or_Control != "Mock")

## order based on library size
expReads <- sampTrack[order(sampTrack$nonchim), ]

## number of rows and start the count from 1 to nrow.
expReads$Index <- seq(nrow(expReads))

## Draw figures
ggplot(data=expReads, aes(x=Index, y=nonchim, color=Sample_Type)) + geom_point() +
  labs(y="Library Size", x="Index Number") + scale_y_continuous( expand = c(0.05,0), limits = c(0, NA), labels = function(x) format(x, scientific = FALSE)) + coord_cartesian(clip='off')

```


Library size of all samples and controls with the Mocks.


```{r qcFilter, message = FALSE, fig.cap = "Reads library size after reads filtering"}

## Filter Control with 0 reads

usectrl <- expReads %>% dplyr::select(Samples,Sample_Type, nonchim) %>% filter(Sample_Type == "Negative" & nonchim > 0) %>% dplyr::select(Samples)

## Filter samples and Mock with more than params$reads reads
useexp <- expReads %>% dplyr::select(Samples, Sample_Type, nonchim) %>% filter(Sample_Type != "Negative" & nonchim > min_reads ) %>% dplyr::select(Samples) 

if (nrow(useexp) <= 1) {
  print ("Warning: After filtering Mock and Negative out, no sample left. Please recheck your metafile or  modify your min reads to include more samples.")
  #knitr::knit_exit()
} 

## Merge two filtered samples
filterdExp <- rbind(useexp, usectrl)

keepSamp <- filterdExp$Samples

filterdExp <- expReads[expReads$Samples %in% keepSamp,]

## Draw figures
ggplot(data=filterdExp, aes(x=Index, y=nonchim, color=Sample_Type)) + geom_point() + labs(y="Library Size", x="Index Number") + scale_y_continuous( expand = c(0.05,0), limits = c(0, NA), labels = function(x) format(x, scientific = FALSE)) + coord_cartesian(clip='off')

write.csv(filterdExp, file.path(reportDir, "filterdExp.csv"))
```


After the minimum reads filtering stage only `r nrow(filterdExp)` samples and controls were left.

