---
title: "Mock taxonomy"
output: html_document

---


```{r mockLib, include = FALSE}

library(plyr)
library(tidyverse)
library(microbiome)
library(phyloseq)

```


## Mock analysis


Depending on the testing kit the relative abundance of the Mock might be different. However, most of the time the frequency should be equal across all 8 genus: 
*Staphylococcus, Salmonella, E.coli, Bacillus, Listeria, Lactobacillus, Enterococcus, Pseudomonas*


```{r mockProcess}

## Keeping only the mocks
mockPs <- prune_samples(sample_data(fltps.ContamFree)$Sample_Type == "Mock", fltps.ContamFree)

## Remove empty taxa
mockPs <- prune_taxa(taxa_sums(mockPs) > 0, mockPs)

## Remove tree
mockPs@phy_tree <- NULL

## Merge identical species
#mockPsSpecies <- tax_glom(mockPs, taxrank="Species")
mockPsSpecies <- tax_glom(mockPs, taxrank="Genus")

## Turn to relative abundance
mockPsSpecies  = transform_sample_counts(mockPsSpecies, function(x) x / sum(x) )

## Add sciName 
mockTax <- tax_table(mockPsSpecies)
mockTaxdf <- data.frame(mockTax)

#mockTaxdf <- mockTaxdf %>% mutate(sciName = paste0(gsub("g__","",Genus), gsub("s_","",Species)))
mockTaxdf <- mockTaxdf %>% mutate(sciName = paste0(gsub("g__","",Genus)))
tax_table(mockPsSpecies) <- as.matrix(mockTaxdf)

# Create dataframe from phyloseq object
mockdf <- psmelt(mockPsSpecies)

# Turn sciName to a character vector from a factor for R
mockdf$sciName <- as.character(mockdf$sciName)

# Group dataframe by Species, calculate median relative abundance
medians <- ddply(mockdf, ~sciName, function(x) c(median=median(x$Abundance)))

# Find Phyla whose rel. abund. is less than 1%
remainder <- medians[medians$median <= 0.01,]$sciName

## Condition if no genus below threshold
if (!identical(character(0), remainder)) {

  # Change their name to "Others"
  mockdf[mockdf$sciName %in% remainder,]$sciName <- 'Others'
}

ggplot(mockdf, aes(x = Sample, y = Abundance, fill = sciName)) + 
    geom_bar(stat = "identity") +
#    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
    coord_flip()

mockFound <- length(medians[medians$median > 0.01,]$sciName)

mockcsv <- mockdf[mockdf$Abundance > 0, c("OTU","Sample","Abundance","sciName")]
write.csv(mockcsv, file.path(reportDir,"mockTaxa.csv"))

```

`r if (mockAnalysis) {paste0("There are ", mockFound, " main taxa with the full taxonomy file can be downloaded: ",  xfun::embed_file( file.path(reportDir,"mockTaxa.csv"), text = "Here !")) }`


