---
title: "Pre sample processing"
output: html_document

params:
  projectname:
    label: "Name of the report"
    value: "MockTest"
    input: text
  dirRaw:
    label: "Working directory"
    value: "../raw"
    input: text
  dirOut:
    label: "Output directory"
    value: "../"
    input: text
  taxo:
    label: "Typing method: dada2 (species) or DECIPHER (genus)"
    input: radio
    value: DECIPHER
    choices: ["dada2", "DECIPHER"]
  decontamMethod:
    label: "Decontam method"
    input: radio
    value: either
    choices: [ "frequency", "prevalence","combined","minimum","either","both" ]
  taxoLevel:
    label: "Agglomerate taxonomy"
    input: radio
    value: family
    choices: ["phylum", "class", "order", "family", "genus", "species"]
    
---

```{r setup3, echo = FALSE}

knitr::opts_chunk$set(echo = FALSE, fig.align="center")

```


```{r reproduce3}

## Change to FALSE if you ran the analysis and only want to produce the report again
firstRun3 <- TRUE

```


```{r lib3, include = FALSE}

library("tidyverse")
library("phyloseq")
library("decontam")
library("microbiome")
library("DECIPHER")
library("phangorn")
library("gridExtra")
library("Biostrings")

start_time <- Sys.time()

```


```{r paramDirs3, child = "conditions/0dirCheck.Rmd"}

```


```{r taxoParams}

## Check experimental file from main script
## If the parameter is empty turns on standalone settings
if (is.null(params$taxa)) {

  typeTyping <- "dada2"
} else {

  typeTyping <- "dada2"
}

deciTaxid <- file.path(reportDir, "taxaDecipher.RDS")
dadaTaxid <- file.path(reportDir, "taxaDada.RDS")

if (typeTyping == "dada2") {
  
  if(file.exists(dadaTaxid)){

    taxid <- readRDS(dadaTaxid)
    dadaTyping = TRUE
  } else {

    print ("Missing dada2 taxonomy file. Please go back one stage.")
    knitr::knit_exit()
  }
} else if (typeTyping == "DECIPHER") {
  
  if(file.exists(deciTaxid)){

    taxid <- readRDS(deciTaxid)
    dadaTyping = FALSE
  } else {

    print ("Missing Decipher taxonomy file. Please go back one stage.")
    knitr::knit_exit()
  }
}

if(file.exists(file.path(reportDir, "filterdExp.csv"))){

  filterdExp <- read.csv(file.path(reportDir, "filterdExp.csv"))
  filterdExp$X <- NULL
} else {

  print ("Missing filtered experiment file. Please go back one stage.")
  knitr::knit_exit()
}

if(file.exists(file.path(reportDir, paste0(projectn, "_seqtab_fin.RDS")))){

  seqtab.fin <- readRDS(file.path(reportDir, paste0(projectn, "_seqtab_fin.RDS")))
} else {

  print ("Missing final seqtab file. Please go back one stage.")
  knitr::knit_exit()
}

if(is.null(params$deconMethod)){

  methdContam <- params$decontamMethod
} else {

  methdContam <- deContam
}

```


```{r taxoDF}

## Turn into a dataframe
dftaxid <- as.data.frame(unlist(taxid))

```


```{r taxDada2Genus, eval = dadaTyping}

dftaxid <- dftaxid %>% rownames_to_column("Seq") %>% dplyr::select(Seq,Kingdom,Phylum,Class,Order,Family,Genus,Species) %>% mutate(Kingdom=paste0("k__",Kingdom), Phylum=paste0("p__",Phylum), Class=paste0("c__",Class), Order=paste0("o__", gsub(" ","_",Order)), Family=paste0("f__",gsub(" ","_",Family)), Genus=paste0("g__", gsub(" ","_",Genus)), Species=paste0("s__", gsub(" ","_",Species)))

## Join taxa together
pastecols <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")

## Taxonomy column
#dftaxid$taxonomy <- apply( dftaxid[ , pastecols ] , 1 , paste , collapse = ";" )

## Reordering columns
#dftaxid <- dftaxid %>% dplyr::select(Seq,taxonomy,Kingdom,Phylum,Class,Order,Family,Genus,Species)

## Change Seq to rownames
rownames(dftaxid) <- dftaxid$Seq
dftaxid$Seq <- NULL

```


```{r taxDecipherGenus, eval = !(dadaTyping) }

## Taxa at genus level
dftaxid <- dftaxid %>% rownames_to_column("Seq") %>% dplyr::select(Seq,Kingdom,Phylum,Class,Order,Family,Genus) %>% mutate(Kingdom=paste0("k__",Kingdom), Phylum=paste0("p__",Phylum), Class=paste0("c__",Class), Order=paste0("o__", gsub(" ","_",Order)), Family=paste0("f__",gsub(" ","_",Family)), Genus=paste0("g__", gsub(" ","_",Genus)))

## Join taxa together
pastecols <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")

## Taxonomy column
#dftaxid$taxonomy <- apply( dftaxid[ , pstcols ] , 1 , paste , collapse = ";" )

## Reordering columns
#dftaxid <- dftaxid %>% dplyr::select(Seq,taxonomy,Kingdom,Phylum,Class,Order,Family,Genus)

## Change Seq to rownames
rownames(dftaxid) <- dftaxid$Seq
dftaxid$Seq <- NULL

```


```{r datataxid}

write.csv(dftaxid, file.path(reportDir,"dftaxid.csv"))

```


```{r alignDNA, eval = firstRun3, included = FALSE, warning = FALSE}

## Gets DNA sequences
seqs <- rownames(dftaxid)
names(seqs) <- seqs

## Align DNA seqs
alignment <- AlignSeqs(DNAStringSet(seqs), anchor=NA,verbose=FALSE)

phangAlign <- phyDat(as(alignment, "matrix"), type="DNA")
dm <- dist.ml(phangAlign)
treeNJ <- NJ(dm) # Note, tip order != sequence order

## pml: computes the likelihood of a phylogenetic tree given a sequence alignment and a model
fit = pml(treeNJ, data=phangAlign)

## update allows to change parameters
## use njt to fit GTR+G+I (Generalized time-reversible with Gamma rate variation) maximum likelihood tree
fitGTR <- update(fit, k=4, inv=0.2)

## Optimize all the parameters.
fitGTR <- optim.pml(fitGTR, model="GTR", optInv=TRUE, optGamma=TRUE,
        rearrangement = "stochastic", control = pml.control(trace = 0))
#detach("package:phangorn", unload=TRUE)

write.tree(fitGTR$tree,file.path(reportDir,"rawTree.tre"))
saveRDS(fitGTR, file.path(reportDir,"rawTree.RDS"))

```


```{r treeCheck}

if (!file.exists(file.path(reportDir,"rawTree.RDS"))) {

  print ("Missing tree file. Please rerun")
  knitr::knit_exit()
}

```


## Taxonomic classification 


The phyloseq file:


```{r phyloseq}

dftaxid <- read.csv(file.path(reportDir,"dftaxid.csv"))
rownames(dftaxid) <- dftaxid$X
dftaxid$X <- NULL
#dftaxid$taxonomy <- NULL

## Load in raw tree
fitGTR <- readRDS(file.path(reportDir,"rawTree.RDS"))
ps_tree <- phy_tree(fitGTR$tree)

filterdExp <- read.csv(file.path(reportDir, "filterdExp.csv"))

## Turn col Row.names to rownames
rownames(filterdExp) <- filterdExp$Samples
filterdExp$Samples <- NULL
filterdExp$Index <- NULL
filterdExp$nonchim <- NULL
filterdExp$readLength <- NULL
filterdExp$X <- NULL

ps <- phyloseq(otu_table(seqtab.fin, taxa_are_rows=FALSE),
               sample_data(filterdExp),
               tax_table(as.matrix(dftaxid)), ps_tree)

## Gets DNA sequences
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))

## Number of taxonomies
tax_found <- ntaxa(ps)

## Filter taxa with no samples
fltps <- prune_taxa(taxa_sums(ps) > 0, ps)

## Filter samples with no reads (Condition for custom run)
fltps <- prune_samples(sample_sums(fltps) > 0, fltps)

taxFlt <- ntaxa(fltps)
nsampFlt <- nsamples(fltps)

saveRDS(fltps, file.path(reportDir,"ps_no_Chime.RDS"))

print (fltps)

```


All the reads are merged together and were subject of classification using using `r typeTyping`. There are `r taxFlt` taxa in `r nsampFlt` samples. The taxa abundance was calculated by the sum per-taxon. The phyloseq file can be downloaded: `r xfun::embed_file(file.path(reportDir,"ps_no_Chime.RDS"), text = "Here!")`


```{r taxaAbundFigRaw, warning = FALSE}

fltps <- readRDS(file.path(reportDir,"ps_no_Chime.RDS"))

### Draws the Phylum abundance figure with everything
## Compute number of sample for each ASV, store as data.frame
abunfltdf = apply(X = otu_table(fltps),
                  MARGIN = ifelse(taxa_are_rows(fltps), yes = 1, no = 2),
                  FUN = function(x){sum(x > 0)})

## Add taxonomy and total read counts to this data.frame
abunfltdf = data.frame(Prevalence = abunfltdf,
                       TotalAbundance = taxa_sums(fltps),
                       tax_table(fltps)[, "Phylum"])

ggplot(abunfltdf, aes(TotalAbundance, Prevalence / nsamples(fltps),color=Phylum)) +
   geom_point(size = 1, alpha = 0.5) +
   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.3)) +
   scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
   facet_wrap(~Phylum) + theme(legend.position="none")

```


For individual sample, each ASV’s counts were divided by the total number of counts to get proportion. Each dot presents an ASV at the Phylum level. Prevalence shows how often such ASV appears in all samples. Total abundance displays the number of reads per ASV.    


```{r phyFeatures}

phyTable <- table(tax_table(fltps)[, "Phylum"], exclude = NULL) %>% as.data.frame()
names(phyTable) <- c("Phylum","Features")
phyTable <- phyTable %>% arrange(desc(Features))

knitr::kable(phyTable, caption = "Number of ASVs/features for each phyla")

```


```{r decontamCheck}

colTitles <- c("DNA_con","Sample_Type")
contamRemov = FALSE

## Condition for decontam prevalence or frequency methods 
if ((any(colTitles %in% colnames(filterdExp))) && ( !("" %in% unique(filterdExp$DNA_con)))) {

  contamRemov = TRUE
}


```


```{r decontam, child = if (contamRemov) "conditions/3decontam.Rmd", warning = FALSE}

```


```{r afterDecontam}

if (contamRemov) {

  ## Input is contamfree phyloseq
  fltps.ContamFree <- readRDS(file.path(reportDir, "fltpsContamFree.RDS"))
  ## Remove the negative column
  sample_data(fltps.ContamFree)$is.neg <- NULL
} else {
  
  ## No chimera ps as decontamination method was not provided 
  fltps.ContamFree <- readRDS(file.path(reportDir,"ps_no_Chime.RDS"))
}

## Remove taxa with 0 count
fltps.ContamFree <- prune_taxa(taxa_sums(fltps.ContamFree) > 0, fltps.ContamFree)

```


```{r finalReads}

if (file.exists(file.path(reportDir,"reads_tracking.csv"))) {
  track <- read.csv(file.path(reportDir,"reads_tracking.csv"))
  track$X <- NULL

  if (contamRemov) {

    track$final_reads <- NULL
    fin_reads <- data.frame(otu_table(fltps.ContamFree))
    fin_reads$final_reads <-  apply(fin_reads, 1, FUN=function(x) sum(x))
    fin_reads <- fin_reads %>% rownames_to_column("Samples")

    track <- merge(track, fin_reads[,c("Samples","final_reads")], by="Samples")

    write.csv(track, file.path(reportDir,"reads_tracking.csv"))
    knitr::kable(head(track), caption = "Tracking reads though the preprocessing steps")
  } else {

    knitr::kable(head(track), caption = "Tracking reads though the preprocessing steps")
  }
}

```


For the complete data of all samples please download  `r if (file.exists(file.path(reportDir,"reads_tracking.csv"))) { xfun::embed_file( file.path(reportDir,"reads_tracking.csv"), text = "Here !") }`


```{r mockCheck}

mockAnalysis <- FALSE
if ("Mock" %in% unique(sample_data(fltps.ContamFree)$Sample_Type)) {

  mockAnalysis <- TRUE
}

```


```{r mockInfo, child = if (mockAnalysis) "conditions/3mockinfo.Rmd"}

```


## Final files


Mocks and Negative controls are were removed from the final phyloseq file. The final phylum composition is shown below.


```{r phyFinFeatures, eval = mockAnalysis}

## Remove Mock samples
finPs <- prune_samples(sample_data(fltps.ContamFree)$Sample_Type != "Mock", fltps.ContamFree)

if (any(sample_data(finPs)$Sample_Type != "Negative")) {

  finPs <- prune_samples(sample_data(finPs)$Sample_Type != "Negative", finPs)
}

```


```{r phyFeature, eval = !(mockAnalysis)}

finPs <- fltps.ContamFree

```


```{r finalPhyloseq}

## Remove empty taxa
finPs <- prune_taxa(taxa_sums(finPs) > 0, finPs)

print (finPs)
saveRDS(finPs, file.path(reportDir, paste0(projectn,"_finPS.rds")))

### Draws the final Phylum abundance figure
abunfltdf2 = apply(X = otu_table(finPs),
                  MARGIN = ifelse(taxa_are_rows(finPs), yes = 1, no = 2),
                  FUN = function(x){sum(x > 0)})

## Add taxonomy and total read counts to this data.frame
abunfltdf2 = data.frame(Prevalence = abunfltdf2,
                       TotalAbundance = taxa_sums(finPs),
                       tax_table(finPs)[, "Phylum"])

ggplot(abunfltdf2, aes(TotalAbundance, Prevalence / nsamples(finPs),color=Phylum)) +
   geom_point(size = 1, alpha = 0.5) +
   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.3)) +
   scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
   facet_wrap(~Phylum) + theme(legend.position="none")

phyTable <- table(tax_table(finPs)[, "Phylum"], exclude = NULL) %>% as.data.frame()
names(phyTable) <- c("Phylum","Features")
phyTable <- phyTable %>% arrange(desc(Features))

knitr::kable(phyTable, caption = "Number of ASVs/features for each phyla")

```


```{r treefig}

taxLevel <- "Genus"
finpsCustom <- tax_glom(finPs, taxrank=taxLevel, NArm = TRUE)

multiPlotTitleTextSize = 15
p1tree = plot_tree(finPs, method = "treeonly",
                   ladderize = "left",
                   title = "Before agglomeration") +
  theme(plot.title = element_text(size = multiPlotTitleTextSize))

p2tree = plot_tree(finpsCustom, method = "treeonly",
                   ladderize = "left", title = paste0("By ",taxLevel)) +
   theme(plot.title = element_text(size = multiPlotTitleTextSize))

grid.arrange(nrow = 1, p1tree, p2tree)

## Gets abundance table for each sample
#treSamp = as.data.frame(microbiome::abundances(fltps.ContamFree))
#treSamp$ASVs <- row.names(treSamp)

## Get taxonomy out of phyloseq
#treTax <- as.data.frame(tax_table(fltps.ContamFree))
#treTax$ASVs <- row.names(treTax)

```


There are `r length(get_taxa_unique(finPs, taxonomic.rank = taxLevel))` unique taxa at this level.


```{r finalextract}

asvTable <- otu_table(finPs)
write.csv(asvTable, file.path(reportDir,"finASV.csv"))

taxaTble <- tax_table(finPs)
write.csv(taxaTble, file.path(reportDir,"finTAXA.csv"))

sampls <- sample_data(finPs)
write.csv(sampls, file.path(reportDir,"finSAMP.csv"))

treePhy <- phy_tree(finPs)
saveRDS(treePhy, file.path(reportDir,"finTree.RDS"))


```


The final phyloseq can be downloaded: `r xfun::embed_file(file.path(reportDir, paste0(projectn,"_finPS.rds")), text = "Here")`.
Separate files include: `r xfun::embed_file(file.path(reportDir,"finASV.csv"), text = "ASVs")`, `r xfun::embed_file(file.path(reportDir,"finTAXA.csv"), text = "Taxa")`, `r xfun::embed_file(file.path(reportDir,"finSAMP.csv"), text = "Experiments")` and `r xfun::embed_file(file.path(reportDir,"finTree.RDS"), text = "Tree")`.


```{r end3}

end_time <- Sys.time()
end_time - start_time

```
